Bias correction in GLMs

### 2021-06-26 bias_correction.rmd
thinksync
Not clear yet!

```{R simple_vareffects, fig.cap = "Variable prediction plots comparing bias unadjusted and adjusted predictions. The bias adjusted predictions are slighly lower but close to what we expect from effect size estimates (and their confidence intervals)?"}
# Age
## Not bias adjusted
simple_vareff_age <- varpred(simple_mod, "age", isolate=TRUE, modelname="Not adjusted")
print(sigma(simple_mod))

## Bias adjusted
simple_vareff_age_adjust <- varpred(simple_mod, "age", isolate=TRUE, bias.adjust=TRUE, modelname="Bias adjusted")

vareff_age <- simple_vareff_age
vareff_age$preds <- do.call("rbind", list(vareff_age$preds, simple_vareff_age_adjust$preds))
age_plot <- (plot(vareff_age)
 	+ labs(y="Prob. of improved \n service", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ theme(legend.position="bottom")
)

# Wealth index
## Not bias adjusted
simple_vareff_wealthindex <- varpred(simple_mod, "wealthindex", isolate=TRUE, modelname="Not adjusted")

## Bias adjusted
simple_vareff_wealthindex_adjust <- varpred(simple_mod, "wealthindex", isolate=TRUE, bias.adjust=TRUE, modelname="Bias adjusted")

vareff_wealthindex <- simple_vareff_wealthindex
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, simple_vareff_wealthindex_adjust$preds))
wealthindex_plot <- (plot(vareff_wealthindex)
 	+ labs(y="", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ theme(legend.position="bottom")
)
ggarrange(age_plot, wealthindex_plot, common.legend=TRUE)
```

@@

\subsection{MRP approach}

```{R simple_mrp}
pred <- predict(simple_mod, type="link", se.fit=TRUE)
z.val <- qnorm(1 - (1 - 0.95)/2)
pred_df <- (data.frame(age=sim_df$age, wealthindex=sim_df$wealthindex, estimate=pred$fit, se=pred$se.fit)
	%>% mutate(lwr = plogis(estimate-z.val*se)
		, upr = plogis(estimate+z.val*se)
		, estimate = plogis(estimate)
	)
)
head(pred_df)

## Age plot
age_mrp <- (pred_df
	%>% select(age, estimate, lwr, upr)
	%>% group_by(age)
	%>% summarise_all(mean)
)
age_mrp_plot <- (ggplot(age_mrp, aes(x=age, y=estimate))
	+ geom_line()
	+ geom_line(aes(y=lwr), lty=2)
	+ geom_line(aes(y=upr), lty=2)
#	+ geom_smooth(aes(ymin=lwr, ymax=upr), stat="identity")
)
print(age_mrp_plot)
```

>>>>

```{R simple_vareffects} 
# Age
## varpred way
simple_vareff_age <- varpred(simple_mod, "age", isolate=FALSE, modelname="varpred")

# Wealth index
## Not bias adjusted
simple_vareff_wealthindex <- varpred(simple_mod, "wealthindex", isolate=TRUE, modelname="varpred")

## Bias adjusted
simple_vareff_wealthindex_adjust <- varpred(simple_mod, "wealthindex", isolate=TRUE, pop.ave=TRUE, modelname="Pop. ave")

vareff_wealthindex <- simple_vareff_wealthindex
vareff_wealthindex$preds <- do.call("rbind", list(vareff_wealthindex$preds, simple_vareff_wealthindex_adjust$preds))
wealthindex_plot <- (plot(vareff_wealthindex)
 	+ labs(y="", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ theme(legend.position="bottom")
)
```

@@

\subsection{Population averaging}

- Averages of the entire population of the non-focal predictor

```{R pop_ave, fig.cap = "A comparison of population averaged and varpred-based predictions. For `age`, we implement the naive approach to compute the predictions in `popavefun` function and then implement the same in `vareffects` so as to use the centering machineries. In both cases, the population averaging and varpred gives similar estimates. The estimated population average is very close to the observed in the case of `wealthindex` but slightly higher in the case of `age`, see previous paragraph."}

popavefun <- function(mod, focal, non.focal, level=0.95, modelname="Pop. ave", ...) {
	mf <- model.matrix(mod)
	mm <- (mf
		%>% data.frame()
		%>% mutate_at(non.focal, mean)
		%>% as.matrix()
	)
	vc <- vcov(mod)
	linpred <- as.vector(mm %*% coef(mod))
	pse_var <- sqrt(rowSums(mm * t(tcrossprod(data.matrix(vc), mm))))
	z.val <- qnorm(1 - (1 - level)/2)
	pred_df <- (mf
		%>% data.frame()
		%>% select_at(focal)
		%>% mutate(fit = linpred
			, lwr = plogis(fit - z.val*pse_var)
			, upr = plogis(fit + z.val*pse_var)
			, fit = plogis(fit)
			, model = modelname
			, se = NA
		)
	)
	return(pred_df)
}

# simple_vareff_age_pop <- varpred(simple_mod, "age", isolate=TRUE, modelname="Pop. ave")
simple_vareff_age_pop <- popavefun(simple_mod, "age", "wealthindex", modelname = "Pop. ave")

vareff_age <- simple_vareff_age
vareff_age$preds <- do.call("rbind", list(vareff_age$preds, simple_vareff_age_pop))
age_plot <- (plot(vareff_age)
 	+ labs(y="Prob. of improved \n service", colour="Model")
	+ geom_hline(yintercept=true_prop, lty=2, colour="grey")
	+ theme(legend.position="bottom")
)

ggarrange(age_plot, wealthindex_plot, common.legend=TRUE)
```

The observed population average is `r true_prop` while the estimated population averages are:

- age: `r mean(simple_vareff_age_pop$fit)`
- wealthindex: `r mean(simple_vareff_wealthindex_adjust$preds$fit)`
